---
title: "Analysis of crude rates of symptomatic and confirm cases"
runtime: shiny
output:
  html_document:
    df_print: paged
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(summarytools)
st_options(
  headings = FALSE,
  footnote = NA
)
source("../.Rprofile")
devtools::load_all("..")
extract_data()
load_data()
library(ggplot2)
library(plotly)
library(shiny)
ch_data <- table_weekly_cases_home()

```


This report is based on `r formatbm(nrow(incidents))` Datix reports of incidents dated up to `r format(max(incidents$incident_date), "%A %e %b %Y")`. Incidents relate to `r formatbm(dplyr::n_distinct(incidents$resident_encryptedid))` unique residents. 

Incidents report were obtained from `r formatbm(dplyr::n_distinct(incidents$home_code))` care homes totalling `r formatbm(sum(beds[beds$year=="2020" & beds$home_code %in% unique(incidents$home_code), "beds"]))` beds. In 2020, Four Seasons Health Care comprises `r formatbm(dplyr::n_distinct(beds[beds$year=="2020", "home_code"]))` care homes with a total of `r formatbm(sum(beds[beds$year=="2020", "beds"]))` beds.

 

```{r, echo=FALSE, fig.cap="Funnel plots of rates of symptomatic and confirmed cases with 95% (blue) and 99% (orange) Poisson control limits"}


shinyApp(
  
  ui = fluidPage(
    sliderInput("date_range", "Date range:",
                min = min(ch_data$week_starting, na.rm = TRUE), 
                max = max(ch_data$week_starting, na.rm = TRUE),
                step = 7,
                width = "100%",
                value = as.Date(c("2020-05-11", "2020-05-25")),
                timeFormat = "%Y-%m-%d"), 
    h5("Funnel plots with 95% (blue) and 99% (orange) Poisson control limits"),
    fluidRow(
      column(width = 6, plotlyOutput("funnelPlot1")),
      column(width = 6, plotlyOutput("funnelPlot2"))
    ), br(),
    fluidRow(
      column(width = 6, plotlyOutput("funnelPlot3")),
      column(width = 6, plotlyOutput("funnelPlot4"))
    ), br(),
    h5("Comparison of Datix and manager counts"),
    fluidRow(
      column(width = 6, plotlyOutput("validation_symptomatic")),
      column(width = 6, plotlyOutput("validation_confirmed"))
    )
   
  ),
  
  server = function(input, output) {
    
    t <- list(
      family = "arial, sans-serif",
      size = 12 )
    
    output$funnelPlot1 = renderPlotly({
      temp_funnel <- ch_data %>% 
        dplyr::filter( week_starting >= input$date_range[1] & 
                         week_starting < input$date_range[2]) %>% 
        plot_funnel(y = first_symptomatic, denom = occupancy_imputed,
                    home_code = home_code, home_name = home_name_clickview) 
      ggplotly(temp_funnel, tooltip = c("text", "rate_per_100000", "observed_cases", "denom")) %>% 
        plotly::layout(title = "Symptomatic cases (Datix)", font = t)
    })
    output$funnelPlot2 = renderPlotly({
      temp_funnel <- ch_data %>% 
        dplyr::filter( week_starting >= input$date_range[1] & 
                         week_starting < input$date_range[2]) %>% 
        plot_funnel(y = tally_first_symptomatic, denom = occupancy_imputed,
                    home_code = home_code, home_name = home_name_clickview)
      ggplotly(temp_funnel, tooltip = c("text", "rate_per_100000", "observed_cases", "denom")) %>% 
        plotly::layout(title = "Symptomatic cases (manager tallies)", font = t)
    })
    output$funnelPlot3 = renderPlotly({
      temp_funnel <- ch_data %>% 
        dplyr::filter( week_starting >= input$date_range[1] & 
                         week_starting < input$date_range[2]) %>% 
        plot_funnel(y = first_confirmed, denom = occupancy_imputed,
                    home_code = home_code, home_name = home_name_clickview)
      ggplotly(temp_funnel, tooltip = c("text", "rate_per_100000", "observed_cases", "denom")) %>% 
        plotly::layout(title = "Confirmed cases (Datix)", font = t)
    })
    output$funnelPlot4 = renderPlotly({
      temp_funnel <- ch_data %>% 
        dplyr::filter( week_starting >= input$date_range[1] & 
                         week_starting < input$date_range[2]) %>% 
        plot_funnel(y = tally_confirmed, denom = occupancy_imputed,
                    home_code = home_code, home_name = home_name_clickview) 
       ggplotly(temp_funnel, tooltip = c("text", "rate_per_100000", "observed_cases", "denom")) %>% 
        plotly::layout(title = "Confirmed cases (manager tallies)", font = t)
    })
    
    output$validation_symptomatic <- renderPlotly({
      temp_data <- ch_data %>% 
        dplyr::filter( week_starting >= input$date_range[1] & 
                         week_starting < input$date_range[2]) %>%
        dplyr::group_by(home_code, home_name_clickview) %>%
        dplyr::summarise(
          tally_first_symptomatic = sum(tally_first_symptomatic, na.rm = T),
          first_symptomatic = sum(first_symptomatic, na.rm = T)
          )

      temp_data <- dplyr::distinct(reference_homes, home_code, home_name_clickview) %>% 
        dplyr::left_join(temp_data)
      temp_data[[3]] <- tidyr::replace_na(temp_data[[3]], 0)
      temp_data[[4]] <- tidyr::replace_na(temp_data[[4]], 0)
      
      temp_scatter <- ggplot(temp_data, aes(x = tally_first_symptomatic, y = first_symptomatic)) + 
        geom_jitter(aes(text = paste("care_home:", home_name_clickview))) +
        geom_abline(aes(intercept = 0, slope = 1,  col = "Identity", linetype = "Identity")) +
        geom_smooth(aes(col="OLS", linetype = "OLS"), method = "lm", se = T) +
        xlab("Manager counts") + ylab("Datix-derived counts") + 
        labs(title="Symptomatic cases") +
        scale_colour_manual(name="Line of fit",
                            values = c("Identity" = "black", "OLS" = "blue", "Spline" = "red")) +
        scale_linetype_manual(name="Line of fit",
                              values = c("Identity" = "solid", "OLS" = "longdash", "Spline" = "dotted")) 
      
      plotly::layout(plotly::ggplotly(temp_scatter), font = t)
    })
    
    output$validation_confirmed <- renderPlotly({
      temp_data <- ch_data %>% 
        dplyr::filter( week_starting >= input$date_range[1] & 
                         week_starting < input$date_range[2]) %>%
        dplyr::group_by(home_code, home_name_clickview) %>%
        dplyr::summarise(
          tally_confirmed = sum(tally_confirmed, na.rm = T),
          first_confirmed = sum(first_confirmed, na.rm = T)
          )

      temp_data <- dplyr::distinct(reference_homes, home_code, home_name_clickview) %>% 
        dplyr::left_join(temp_data)
      temp_data[[3]] <- tidyr::replace_na(temp_data[[3]], 0)
      temp_data[[4]] <- tidyr::replace_na(temp_data[[4]], 0)
      
      temp_scatter <- ggplot(temp_data, aes(x = tally_confirmed, y = first_confirmed)) + 
        geom_jitter(aes(text = paste("care_home:", home_name_clickview))) +
        geom_abline(aes(intercept = 0, slope = 1,  col = "Identity", linetype = "Identity")) +
        geom_smooth(aes(col="OLS", linetype = "OLS"), method = "lm", se = T) +
        xlab("Manager counts") + ylab("Datix-derived counts") + 
        labs(title="Confirmed cases") +
        scale_colour_manual(name="Line of fit",
                            values = c("Identity" = "black", "OLS" = "blue", "Spline" = "red")) +
        scale_linetype_manual(name="Line of fit",
                              values = c("Identity" = "solid", "OLS" = "longdash", "Spline" = "dotted")) 
      
      plotly::layout(plotly::ggplotly(temp_scatter), font = t)
    })
    
    
  },
  
  options = list(height = 1500)
)
```
 

