---
title: "Data quality report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(summarytools)
st_options(
  headings = FALSE,
  footnote = NA
)
source("../.Rprofile")
devtools::load_all("..")
extract_data()
load_data()
```


This report is based on n=`r formatbm(nrow(incidents))` Datix reports of incidents dated up to `r format(max(incidents$incident_date), "%A%e %b %Y")`. Incidents relate to `r formatbm(dplyr::n_distinct(incidents$resident_encryptedid))` unique residents. 

Incidents report were obtained from `r formatbm(dplyr::n_distinct(incidents$home_code))` care homes totalling `r formatbm(sum(beds[beds$year=="2020" & beds$home_code %in% unique(incidents$home_code), "beds"]))` beds. In 2020, Four Seasons Health Care comprises `r formatbm(dplyr::n_distinct(beds[beds$year=="2020", "home_code"]))` care homes with a total of `r formatbm(sum(beds[beds$year=="2020", "beds"]))` beds.


## Classification of cases

The incident classification is not wholly consistent, as illustrated in the tabulation below. Due to this, a reclassification algorithm is used.

```{r }
print(ctable(incidents$infection_confirmed, incidents$infection_covid_19_type, useNA = "always"), method = 'render')
```

### Algorithm

The following algorithm is used to classify symptomatic, suspected, and test-confirmed cases from the Datix reports of COVID-19 infections.


```{r echo=FALSE}
DiagrammeR::grViz("../man/figure/case_classification_1.gv", width=700, height= 500)
```

```{r echo = FALSE}
DiagrammeR::grViz("../man/figure/case_classification_2.gv", width=700, height= 400)
```

### Validation

Data are tabulated below to evaluate the internal consistency of the rules set out in the previous section.

The `infection_result` field provides information on test results which broadly agrees yet sometimes conflicts with the `infection_confirmed` classification of incidents. It is potentially more up to date although it is very sparsely populated.


```{r}
print(ctable(incidents$infection_covid_19_type, incidents$covid_confirmed, useNA = "always"), method = 'render')
```

```{r}
print(ctable(incidents$infection_covid_19_type, incidents$covid_test_result, useNA = "always"), method = 'render')
```

```{r}
print(ctable(incidents$infection_confirmed, incidents$covid_test_result, useNA = "always"), method = 'render')
```

## Data quality

### Demographic information

Resident pseudoidentifiers were absent in `r sum(grepl("^MISSING", incidents$resident_encryptedid))` (`r tbrounding(mean(grepl("^MISSING", incidents$resident_encryptedid))*100)`%) of Datix incident reports. However, the pseudoidentifiers of a further `r sum(incidents$resident_encryptedid[which(!grepl("^MISSING", incidents$resident_encryptedid))] %in% unique(residents$encrypted_id))` reports (`r tbrounding(sum(incidents$resident_encryptedid[which(!grepl("^MISSING", incidents$resident_encryptedid))] %in% unique(residents$encrypted_id))/nrow(incidents)*100)`%) were could not be successfully linked to the residents database. In total, `r sum(is.na(incidents$age) & is.na(incidents$date_of_birth))` reports (`r tbrounding(mean(is.na(incidents$age) & is.na(incidents$date_of_birth))*100)`%) did not have any date of birth or age, and `r sum(is.na(incidents$gender))` reports (`r tbrounding(mean(is.na(incidents$gender))*100)`%) did not have sex recorded. 






